A sample article title

### 2016-06-14 additional_files/tex/distribution-notes.tex
commit after harriet move - too many files to remove manually - will try to find what the problems in the .gitignore is

@@

\title{Curating RNAseq data for mouse and \textit{Eimeria falciformis}} % Title

\author{Emanuel \textsc{Heitlinger}} % Author name

@@

% figure 1
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_mouseNocutoff} % Include the image placeholder.png
\caption{Density distribution of raw transcript counts (log10) for all
  mouse samples in analysis. All samples share the same bimodal
  distribution trend by visual inspection. The first density peak
  occur at $1 - 10^2$ transcripts, \textit{i.e.} transcripts which are
  detected between 1 and 100 times in that sample. The second density
  peak occurs in the range of $10^3 - 10^4$.}
\end{center}
\end{figure}

% figure 2
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouseNocutoff} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for all mouse samples. All samples share the same bimodal
  distribution trend by visual inspection. The first density peak
  occurs at $<1$ transcripts, and in most samples it is more distinct
  than in the non-normalised data in Figure 1. For this normalised
  data, the second density peak occurs at roughly the same count value
  in all samples, close to $10^2$ transcripts.}
\end{center}
\end{figure}

% figure 3
\begin{figure}[h]
\begin{center}
  \includegraphics[width=0.8\textwidth]{distributions_normmouse100} % Include the image placeholder.png
  \caption{Density distribution of normalised transcript counts
    (log10) for mouse samples with a cutoff of 100 (see Methods for
    details). All samples still have a bimodal distribution, however
    less pronounced compared to figures 1 and 2. The first density
    peak occurs at $<1$ transcripts as in Figure 2. The second density
    peak occurs at close to $10^2$ transcripts and did hence not
    change from the normalised data without cutoff (Figure 2).}
\end{center}
\end{figure}

% figure 4
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouse2000} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for mouse samples with a cutoff of 2000 (see Methods for
  details). The bimodality seen at lower cutoff values is strongly
  weakened or not detectable visually in some samples. It is however
  visually still seen in a majority of the samples. The first tendency
  to a density peak has shifted somewhat towards higher transcript
  counts (right) but is still $<1$ transcripts as in Figure 3. The
  second density peak occurs at close to $10^2$ transcripts and did
  hence not change from the normalised data with a lower cutoff
  (Figure 3).}
\end{center}
\end{figure}

% figure 5
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouse5000} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for mouse samples with a cutoff of 5000 (see Methods for
  details). The bimodality is not visually detectable in most samples,
  however it is still seen in \textit{e.g.} samples
  NMRI\_1stInf\_3dpi\_rep1 and NMRI\_2ndInf\_7dpi\_rep2. The major
  density peak remains at a value close to $10^2$ as in all the
  analysed normalised datasets (\textit{e.g.} Figure 2 or 4).}
\end{center}
\end{figure}

% figure 6
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_mouse5000} % Include the image placeholder.png
\caption{Density distribution of raw transcript counts (log10) for
  mouse samples with a cutoff of 5000 (see Methods for details). The
  bimodality is not visually detectable in most samples, however it is
  still seen in \textit{e.g.} samples NMRI\_1stInf\_3dpi\_rep2 and
  NMRI\_2ndInf\_7dpi\_rep2. The major density peak is seen at a value
  over $10^2$, shifting it almost one order of magnitude compared to
  the normalised data with the same cutoff (Figure 5).}
\end{center}
\end{figure}

% figure 7
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouse10000} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for mouse samples with a cutoff of 10000 (see Methods for
  details). The bimodality is not visually detectable in any sample
  apart from a minor indication in sample
  NMRI\_1stInf\_5dpi\_rep2. The major density peak remains at a value
  close to $10^2$ as in all the analysed normalised datasets
  (\textit{e.g.} Figure 2 or 4).}
\end{center}
\end{figure}

@@

% figure 8 
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Ef5} % Include the image placeholder.png
\caption{Density distribution of raw transcript counts (log10) for
  parasite samples with a cutoff of 5 (see Methods for details). Upon
  visual inspection distributions in most samples appear smooth and
  not contradictory to a negative binomial distribution. Density peak
  positions with regards to transcript counts (x-axis) are however at
  different orders of magnitudes between samples. Samples
  NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2, however, do
  not.}
\end{center}
\end{figure}

% figure 9
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Efnorm5} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for parasite samples with a cutoff of 5 (see Methods for
  details). Upon visual inspection distributions in most samples
  appear smooth also in the normalised data. Density peaks are also in
  the same order of magnitude with regards to transcript counts
  (x-axis). Non-negative binomial samples in Figure 8
  (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2) have more
  pronounced spikes in this normalised data. Additionally, samples
  Rag\_1stInf\_5dpi\_rep1 and Rag\_1stInf\_5dpi\_rep2 also have one
  additional peak in the lower range of the distribution, as do some
  other samples, but less pronounced.}
\end{center}
\end{figure}

% figure 10
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Ef50000} % Include the image placeholder.png
\caption{Density distribution of raw transcript counts (log10) for
  parasite samples with a cutoff of 50000 (see Methods for
  details). Upon visual inspection distributions in most samples
  appear smooth. Non-negative binomial samples in Figure 8 and 9
  (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2 have more than
  one density peak also when lowly expressed genes are removed with a
  high threshold (50000). Samples Rag\_1stInf\_5dpi\_rep1 and
  Rag\_1stInf\_5dpi\_rep2, however, seem smoother with only one
  density peak with this threshold.}
\end{center}
\end{figure}

% figure 11
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Efnorm50000} % Include the image placeholder.png
\caption{Density distribution of normalised transcript counts (log10)
  for parasite samples with a cutoff of 50000 (see Methods for
  details). Upon visual inspection distributions in most samples
  appear smooth. Non-negative binomial samples are pronounced bimodal
  in this normalised data (NMRI\_2ndInf\_3dpi\_rep1,
  NMRI\_2ndInf\_3dpi\_rep2, NMRI\_2ndInf\_5dpi\_rep2 and
  NMRI\_2ndInf\_7dpi\_rep2). Samples Rag\_1stInf\_5dpi\_rep1 and
  Rag\_1stInf\_5dpi\_rep2 are also bimodal, however less pronounced.}
\end{center}
\end{figure}

@@

%\begin{enumerate}
%\begin{item}
%  The \emph{atomic weight of an element} is the relative weight of
%  one of its atoms compared to C-12 with a weight of
%  12.0000000$\ldots$, hydrogen with a weight of 1.008, to oxygen with
%  a weight of 16.00. Atomic weight is also the average weight of all
%  the atoms of that element as they occur in nature.
%\end{item}
%\begin{item}
%  The \emph{units of atomic weight} are two-fold, with an identical
%  numerical value. They are g/mole of atoms (or just g/mol) or
%  amu/atom.
%\end{item}
%\begin{item}
%  \emph{Percentage discrepancy} between an accepted (literature)
%  value and an experimental value is
%\begin{equation*}
%\frac{\mathrm{experimental\;result} - \mathrm{accepted\;result}}{\mathrm{accepted\;result}}
%\end{equation*}
%\end{item}
%\end{enumerate}

>>>>

%%\usepackage{pdfpages}

@@

\title{Gene expression distribution in RNAseq data for mouse and \textit{Eimeria falciformis}} % Title

\author{Totta \textsc{Kasemo}} % Author name

@@

% figure 1
\begin{figure}[h]
\begin{center}
  \includegraphics[width=0.8\textwidth]{distributions_mouseNocutoffg} 
  \caption{Density distribution of raw transcript counts (log10) for
    all mouse samples in analysis. All samples share the same bimodal
    distribution trend by visual inspection. The first density peak
    occur at $1 - 10^2$ transcripts, \textit{i.e.} transcripts which
    are detected between 1 and 100 times in that sample. The second
    density peak occurs in the range of $10^3 - 10^4$.}
\end{center}
\end{figure}

% figure 2
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouseNocutoff} 
\caption{Density distribution of normalised transcript counts (log10)
  for all mouse samples. All samples share the same bimodal
  distribution trend by visual inspection. The first density peak
  occurs at $<1$ transcripts, and in most samples it is more distinct
  than in the non-normalised data in Figure 1. For this normalised
  data, the second density peak occurs at roughly the same count value
  in all samples, close to $10^2$ transcripts.}
\end{center}
\end{figure}

% figure 3
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouse100} 
\caption{Density distribution of normalised transcript counts (log10)
  for mouse samples with a cutoff of 100 (see Methods for
  details). All samples still have a bimodal distribution, however
  less pronounced compared to figures 1 and 2. The first density peak
  occurs at $<1$ transcripts as in Figure 2. The second density peak
  occurs at close to $10^2$ transcripts and did hence not change from
  the normalised data without cutoff (Figure 2).}
\end{center}
\end{figure}

% figure 4
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouse2000} 
\caption{Density distribution of normalised transcript counts (log10)
  for mouse samples with a cutoff of 2000 (see Methods for
  details). The bimodality seen at lower cutoff values is strongly
  weakened or not detectable visually in some samples. It is however
  visually still seen in a majority of the samples. The first tendency
  to a density peak has shifted somewhat towards higher transcript
  counts (right) but is still $<1$ transcripts as in Figure 3. The
  second density peak occurs at close to $10^2$ transcripts and did
  hence not change from the normalised data with a lower cutoff
  (Figure 3).}
\end{center}
\end{figure}

% figure 5
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouse5000}
\caption{Density distribution of normalised transcript counts (log10)
  for mouse samples with a cutoff of 5000 (see Methods for
  details). The bimodality is not visually detectable in most samples,
  however it is still seen in \textit{e.g.} samples
  NMRI\_1stInf\_3dpi\_rep1 and NMRI\_2ndInf\_7dpi\_rep2. The major
  density peak remains at a value close to $10^2$ as in all the
  analysed normalised datasets (\textit{e.g.} Figure 2 or 4).}
\end{center}
\end{figure}

% figure 6
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_mouse5000}
\caption{Density distribution of raw transcript counts (log10) for
  mouse samples with a cutoff of 5000 (see Methods for details). The
  bimodality is not visually detectable in most samples, however it is
  still seen in \textit{e.g.} samples NMRI\_1stInf\_3dpi\_rep2 and
  NMRI\_2ndInf\_7dpi\_rep2. The major density peak is seen at a value
  over $10^2$, shifting it almost one order of magnitude compared to
  the normalised data with the same cutoff (Figure 5).}
\end{center}
\end{figure}

% figure 7
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_normmouse10000}
\caption{Density distribution of normalised transcript counts (log10)
  for mouse samples with a cutoff of 10000 (see Methods for
  details). The bimodality is not visually detectable in any sample
  apart from a minor indication in sample
  NMRI\_1stInf\_5dpi\_rep2. The major density peak remains at a value
  close to $10^2$ as in all the analysed normalised datasets
  (\textit{e.g.} Figure 2 or 4).}
\end{center}
\end{figure}

@@

% figure 8 
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Ef5}
\caption{Density distribution of raw transcript counts (log10) for
  parasite samples with a cutoff of 5 (see Methods for details). Upon
  visual inspection distributions in most samples appear smooth and
  not contradictory to a negative binomial distribution. Density peak
  positions with regards to transcript counts (x-axis) are however at
  different orders of magnitudes between samples. Samples
  NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2, however, do
  not.}
\end{center}
\end{figure}

% figure 9
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Efnorm5}
\caption{Density distribution of normalised transcript counts (log10)
  for parasite samples with a cutoff of 5 (see Methods for
  details). Upon visual inspection distributions in most samples
  appear smooth also in the normalised data. Density peaks are also in
  the same order of magnitude with regards to transcript counts
  (x-axis). Non-negative binomial samples in Figure 8
  (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2) have more
  pronounced spikes in this normalised data. Additionally, samples
  Rag\_1stInf\_5dpi\_rep1 and Rag\_1stInf\_5dpi\_rep2 also have one
  additional peak in the lower range of the distribution, as do some
  other samples, but less pronounced.}
\end{center}
\end{figure}

% figure 10
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Ef50000}
\caption{Density distribution of raw transcript counts (log10) for
  parasite samples with a cutoff of 50000 (see Methods for
  details). Upon visual inspection distributions in most samples
  appear smooth. Non-negative binomial samples in Figure 8 and 9
  (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
  NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2 have more than
  one density peak also when lowly expressed genes are removed with a
  high threshold (50000). Samples Rag\_1stInf\_5dpi\_rep1 and
  Rag\_1stInf\_5dpi\_rep2, however, seem smoother with only one
  density peak with this threshold.}
\end{center}
\end{figure}

% figure 11
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\textwidth]{distributions_Efnorm50000}
\caption{Density distribution of normalised transcript counts (log10)
  for parasite samples with a cutoff of 50000 (see Methods for
  details). Upon visual inspection distributions in most samples
  appear smooth. Non-negative binomial samples are pronounced bimodal
  in this normalised data (NMRI\_2ndInf\_3dpi\_rep1,
  NMRI\_2ndInf\_3dpi\_rep2, NMRI\_2ndInf\_5dpi\_rep2 and
  NMRI\_2ndInf\_7dpi\_rep2). Samples Rag\_1stInf\_5dpi\_rep1 and
  Rag\_1stInf\_5dpi\_rep2 are also bimodal, however less pronounced.}
\end{center}
\end{figure}

@@

%\begin{enumerate}
%\begin{item}
%The \emph{atomic weight of an element} is the relative weight of one of its atoms compared to C-12 with a weight of 12.0000000$\ldots$, hydrogen with a weight of 1.008, to oxygen with a weight of 16.00. Atomic weight is also the average weight of all the atoms of that element as they occur in nature.
%\end{item}
%\begin{item}
%The \emph{units of atomic weight} are two-fold, with an identical numerical value. They are g/mole of atoms (or just g/mol) or amu/atom.
%\end{item}
%\begin{item}
%\emph{Percentage discrepancy} between an accepted (literature) value and an experimental value is
%\begin{equation*}
%\frac{\mathrm{experimental\;result} - \mathrm{accepted\;result}}{\mathrm{accepted\;result}}
%\end{equation*}
%\end{item}
%\end{enumerate}

### 2016-06-14 distribution-notes.tex
commit after harriet move - too many files to remove manually - will try to find what the problems in the .gitignore is

Mouse\\
The assumption of a negative binomial distribution underlies most
differential gene expression analyses available. Similar bimodal
expression patterns as in our mouse dataset have been reported
(\textit{e.g.} ....) and suggested to be a result of a group of
non-coding RNAs being sequenced. Since we are not interested in such
RNA species in this project, different cutoff values were tested to
find the lowest possible threshold for removing transcripts and gain a
negative binomial distribution. By visual inspection (again: should we
formalise/quantify this somehow?) we chose 5000 as the cutoff for
further analysis.\\

Parasite\\
The parasite data follows the expected distribution in most samples
(although we have not formally tested it - check mean-var relationship
to do that?). Bimodal or multipeak samples do not show unimodal
distributions even with high cutoffs, \textit{i.e.} many transcripts
required to map to one gene for the reads to be included. The four
most obvious spiky samples are also the samples with fewest reads
sequenced overall, which can explain the uneven distribution (make
table to refer to?). We will exclude these four samples
(NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2,
NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2) from all
further analysis. It is not clear why there is such low coverage in
these samples. Possibilities include poor infection success because of
too low oocyst injection, mouse resistance due to \textit{e.g.}
unknown previous infections or other unknown variability in the
experiment. Because of this insecurity, these samples are also
exculded from the mouse data.\\

>>>>

Mouse\\
The assumption of a negative binomial distribution underlies most differential gene expression analyses available. Similar bimodal expression patterns as in our mouse dataset have been reported (\textit{e.g.} ....) and suggested to be a result of a group of non-coding RNAs being sequenced. Since we are not interested in such RNA species in this project, different cutoff values were tested to find the lowest possible threshold for removing transcripts and gain a negative binomial distribution. By visual inspection (again: should we formalise/quantify this somehow?) we chose 5000 as the cutoff for further analysis.  \\

Parasite\\
The parasite data follows the expected distribution in most samples (although we have not formally tested it - check mean-var relationship to do that?). Bimodal or multipeak samples do not show unimodal distributions even with high cutoffs, \textit{i.e.} many transcripts required to map to one gene for the reads to be included. The four most obvious spiky samples are also the samples with fewest reads sequenced overall, which can explain the uneven distribution (make table to refer to?). We will exclude these four samples (NMRI\_2ndInf\_3dpi\_rep1, NMRI\_2ndInf\_3dpi\_rep2, NMRI\_2ndInf\_5dpi\_rep2 and NMRI\_2ndInf\_7dpi\_rep2) from all further analysis. It is not clear why there is such low coverage in these samples. Possibilities include poor infection success because of too low oocyst injection, mouse resistance due to \textit{e.g.} unknown previous infections or other unknown variability in the experiment. Because of this insecurity, these samples are also exculded from the mouse data. \\

### 2016-06-14 manuscript/tex_bmc_format/manuscript_rnaseq.tex
commit after harriet move - too many files to remove manually - will try to find what the problems in the .gitignore is

  \section*{Results}
Text for this section \ldots
\subsection*{GO term enrichments in heatmap clusters}
The annotations referred to here are in many 
cases inferred from the existence of predicted protein domains and have in some cases been 
verified in e.g. \textit{E. tenella} or other \textit{Eimeria} spp.

\subsubsection*{Preparation for invasion in oocysts}
The mRNA profile in the oocyst stage is mainly determined by highly abundant genes in cluster 4.
Overrepresented GO-term in this cluster are largely determined by genes orthologous to 
peptidases, microneme localized proteins reported to be involved in invasion, 
genes associated with clotting and in protozoans, adhesion, and genes that are annotated 
to be involved in amino acid biosynthesis.
  Aminopeptidase N ('related' annotation) is the annotation for orthologs of three genes with 
abundant mRNAs in oocysts. In humans, this enzyme has been reported to cleave peptides bound to 
major histocompatibility complex, MHC, II (UniProt reference if we want to keep this... but does any
secretion happen from oocysts...? Or is this too far-fetched to be interesting?).

Three genes can be linked to adhesion and through localization the same genes
also are asociated with invasion in the literature. 
A protein domain which is associated with adhesion in protozoans, thrombospondin type 1 
domain-containing protein, was found in our oocyst cluster (cluster 1). 
Thrombospondin type 1 domains have been reported in \textit{E. tenella} microneme localizing 
proteins, MIC. Tomley et al. (Tomley01) describe MIC4 as a protein contatining thrombospondin 
type 1 repeats. In \textit{E. tenella} MIC4 is expressed in sporozoites where it localizes to the 
apical end, in late schizonts and late oocyst stages, when sporozoites are forming.
(Tomley01)
In the protozoan \textit{P. falciparum}, another domain occuring in our data, the sushi domain, 
is reported in the apical sushi protein, ASP. In \textit{P. falciparum} it localizes to the 
micronemes in merozoites but not other stages (OKeeffe05). To our knowledge, the role of sushi 
domain proteins in \textit{Eimeria} spp. oocysts has not yet been investigated. 
  Other protein domains detected via orthologs to mRNAs expressed in oocysts in our data are
Limulus clotting factor C, Coch-5b2 (Cochlin) and Lgl1, LCCL, also known as F5/8 domain. 
LCCL domains have  been associated with discoidin lectin domains, which in turn are involved 
in cell adhesion (Pfam entries for 'LCCL domain' and 'Discoidin domain', May 2016). 
In the slime mold \textit{Dictyostelium discoideum)} the LCCL domain
is part of the mold'd discoidin adhesion protein. In humans, the LCCL domain in annotated as a 
coagulation factor domain. (Pfam entry for 'discoidin domain', May 2016)

\subsubsection*{Amino acid biosynthesis in oocysts}
mRNA of a aminotransferases indicate amino acid biosynthesis in oocysts. 
We also identify D-3-phosphoglycerate dehydrogenase and alanine dehydrogenase, which are enzymes 
contributing to L-serine and L-alanine production, respectively. A putative \textit{Eimeria} spp.
cystathionine beta-synthase, CBS, in this cluster also indicates de novo cysteine production. 
Alkyl sulfatase mRNA is another gene which contributes to overrepresentation of a GO-term in 
mRNAs highly abundant in oocysts. Generally, this enzyme enables an organism to exploit organic 
sulfur to produce and incorporate inorganic sulfur into the amino acids cysteine and methionine, 
when no inorganoc sulfur is available.

@@

Propionate is one of two most abundant small-chain fatty acids 
in the gut along with butyrate. Both fatty acids are largely produced as degradation products from food
by commensal bacteria (Sun13). MmgE/PrpD is important for propionate catabolism in the 
2-methylcitric acid cycle and has been shown to be used by the intestinal intracellular bacterium 
\textit{Salmonella typhimurium} to generate pyruvate (Horswill99).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%% DAY xx   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{cluster 3..........}

>>>>

  \section*{Results}
Text for this section \ldots
\subsection*{GO term enrichments in heatmap clusters}
The annotations referred to here are inferred from orthologs in other \textit{Eimeria} spp. or
in \textit{T. gondii}

\subsubsection*{Preparation for invasion in oocysts}
The mRNA profile in the oocyst stage is mainly determined by highly abundant genes in cluster 4.
Overrepresented GO-terms in this cluster are enriched by ortholog genes to 
peptidases, microneme localized proteins reported to be involved in invasion, 
genes associated with adhesion in protozoans (and with clotting in higher eukaryotes),
and genes that are annotated to be involved in amino acid biosynthesis.
  Aminopeptidase N ('related' annotation) is the reported ortholog for three genes with 
abundant mRNAs in oocysts. In humans, this enzyme has been reported to cleave peptides bound to 
major histocompatibility complex, MHC, II (UniProt reference if we want to keep this... but does any
secretion happen from oocysts...? Or is this too far-fetched to be interesting?).

A Thrombospondin type 1 domain-containing protein ortholog is highly abundant in cluster 4 (high 
abundance in oocysts). Thrombospondin type 1 domains have been reported in \textit{E. tenella} microneme
localizing proteins, MIC, e.g. MIC4 (Tomley01) In \textit{E. tenella} MIC4 mRNA was reported in 
sporozoites where it localizes to the apical end, and in late schizonts and late oocyst stages,
when sporozoites are forming. (Tomley01). For the same gene, the \textit{T. gondii} annotation is 
Sushi domain-containing protein, which is also the ortholog annotation of another gene in this 
cluster. In the related malaria parasite \textit{P. falciparum} the apical sushi 
protein, ASP, (which has a sushi domain) localizes to micronemes in merozoites but not 
other stages (OKeeffe05). 
Limulus clotting factor C, Coch-5b2 (Cochlin) and Lgl1, LCCL, 
(syn. F5/8 domain) domains are associated discoidin lectin domains and thereby with adhesion.
(Pfam entries for 'LCCL domain' and 'Discoidin domain', May 2016). 
Taken together, this indicates that the thrombospondin and sushi-domain genes 
(EfaB_MINUS_4114.g412 and EfaB_PLUS_1425.g183) are involved in sporozoite invasion in 
\textit{E. falciformis} and that the mRNAs are transcribed and available before excystation. 
A role in merozoite re-invasion in \textit{E. falciformis} is not indicated by our data.
The LCCL domain annotation and thrombospondins role in higher eukaryotes also indicates 
that adhesion or preparation for adhesion is important in oocysts. 
We suggest (speculate...?) that the thrombospondin annotated ortholog and the LCCL domain-containing protein 
(EfaB_MINUS_11233.g986) are involved in cell adhesion in \textit{E. falciformis}.
%One examples
%is found in the slime mold \textit{Dictyostelium discoideum)} in which the LCCL domain
%is part of the mold's discoidin adhesion protein. (Pfam entry for 'discoidin domain', May 2016)

\subsubsection*{Amino acid biosynthesis in oocysts}
High abundance of aminotransferase mRNAs indicate amino acid biosynthesis or preparation for 
the same in oocysts (cluster 4). 
We identify D-3-phosphoglycerate dehydrogenase and alanine dehydrogenase orthologs, which 
are enzymes contributing to L-serine and L-alanine production, respectively. A putative 
\textit{Eimeria} spp. cystathionine beta-synthase, CBS, in this cluster also indicates 
de novo cysteine production. Alkyl sulfatase mRNA is also abundant in oocysts. Generally, 
this enzyme enables an organism to exploit organic sulfur to produce and incorporate 
inorganic sulfur into the amino acids cysteine and methionine, when no inorganoc sulfur 
is available.

@@

\subsubsection*{Oocysts contain mRNA for fatty acid catabolism}
MmgE/PrpD is overrepresented in oocysts. The enzyme is important for propionate catabolism 
in the 2-methylcitric acid cycle and has been shown to be used by the 
intestinal intracellular bacterium \textit{Salmonella typhimurium} to generate 
pyruvate (Horswill99). 
Propionate is one of two most abundant small-chain fatty acids 
in the gut along with butyrate. Both fatty acids are largely produced as 
degradation products from food by commensal bacteria (Sun13). Sharing the intestines as a niche
with \textit{S. typhimurium} it is possible that also \textit{E. falciformis} uses Mmg/PrpD
to exploit available propionate for pyruvate production.

\subsubsection*{Oocyst highly abundant mRNAs are downregulated in sporozoites}
Interestingly, the genes described above which are thought to be involved in amino acid 
biosynthesis and invasion are highly abundant in oocysts but are underrepresented in schizont
stages (day 3 and day 5 samples) and even in sporozoites. An average abundance was detected on 
day 7 for these genes, indicating a role in either gametes or early oocyst formation.
This pattern supports the suggestion that these specific mRNAs (cluster 4) for invasion 
and biosynthetic processes are prepared (and possibly expressed) in the oocyst stage but 
are no longer detectable in the cell at the timepont when the protein is assumed to be 
in use (sporozoites and merozoite stages). 
Therefore, correlating mRNA prevalence with biological function at the timepoint when mRNAs 
are detected must be done with care.

\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
Oocysts: profile for 3, 5, 6:

Sporozoites: 3 and 5

Day 7: 5 and 7

Specific genes....
Enolase 2, encoded by Eno2, is among the downregulated genes in oocysts and sporozoites.
In \textit{T. gondii} the paralog Eno1 is strongly associated with the cyst (bradyzoite) stage
and Eno2 is associated with tachyzoite stages (kibe05). It is therefore expected that this mRNA is
underrepresented in oocysts and our data also show that the same is true in sporozoites for
\textit{E. falciformis}. (TK: If important we could look specifically for Eno1 in cluster 4).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%% DAY xx   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Down in sporozoites and oocysts -->  cluster 3..........}
